### 2019-11-01 preproc.Rmd

---
title: "Main1"
author: "Yigit Ozan Berk"
date: "11/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data import

```{r}
install.packages("tm")
install.packages("slam")
library(slam)
library(tm)

```

```{r}
con <- file("./en_US.twitter.txt", "r") 
twit = readLines(con, skipNul = T)
close(con)

```

```{r}
set.seed(1)
x = rbinom(2360148, size = 1, prob = .7)
#70% of data for training, 30% for testing
train = twit[x]
test = twit[!x]
rm(twit)
```