### 2019-11-01 preproc.Rmd
week 1 quiz done

---
title: "Main1"
author: "Yigit Ozan Berk"
date: "11/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data import

```{r}
install.packages("tm")
install.packages("slam")
library(slam)
library(tm)

```

## twitter files
```{r}
con <- file("./en_US.twitter.txt", "r") 
twit = readLines(con, skipNul = T)
close(con)
y = NULL
for(i in 1:length(twit)) {y[i] = nchar(twit[[i]])}
summary(y)
```

```{r}
set.seed(1)
x = rbinom(2360148, size = 1, prob = .7)
#70% of data for training, 30% for testing
train = twit[x]
test = twit[!x]
rm(twit)
```

## blogs

```{r}
con <- file("./en_US.blogs.txt", "r") 
blogs = readLines(con, skipNul = T)
close(con)

y = NULL
for(i in 1:length(blogs)) {y[i] = nchar(blogs[[i]])}
summary(y)

set.seed(123)
x = rbinom(899288, size = 1, prob = .7)
#70% of data for training, 30% for testing
train = blog[x]
test = blog[!x]
```