### 2019-11-05 preproc2.Rmd

---
title: "PreProc2"
author: "Yigit Ozan Berk"
date: "11/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# init

```{r, init}
library(quanteda)
library(readtext)
library(spacyr)
```

```{r, numwords}

##orig.wd <- getwd()
##setwd("../final/en_US")
numwords <- system("wc -w *.txt", intern=TRUE)  # intern=TRUE to return output  
numlines <- system("wc -l *.txt", intern=TRUE)
numbytes <- system("wc -c *.txt", intern=TRUE)
# longest <- system("wc -L *.txt", intern=TRUE) # not working
##setwd(orig.wd)  # return to original working dir, ie. the parent of /final

# number of words for each dataset
blog.numwords <- as.numeric(gsub('[^0-9]', '', numwords[1]))
news.numwords <- as.numeric(gsub('[^0-9]', '', numwords[2]))
twit.numwords <- as.numeric(gsub('[^0-9]', '', numwords[3]))

# number of lines for each dataset
blog.numlines <- as.numeric(gsub('[^0-9]', '', numlines[1]))
news.numlines <- as.numeric(gsub('[^0-9]', '', numlines[2]))
twit.numlines <- as.numeric(gsub('[^0-9]', '', numlines[3]))

# number of bytes for each dataset
blog.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[1]))
news.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[2]))
twit.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[3]))
```

# data import
```{r}
dat.twit <- readtext("./en_US.twitter.txt", cache = FALSE)
# summary(corpus(dat.twit), 5)
# or

con <- file("./en_US.twitter.txt", "r") 
twit = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(12345)
x = sample(2360148, 350000, replace = F)

train = twit[x]

myN <- corpus(train)
rm(con, train, twit, x)
summary(myN, 5)
```