### 2019-11-09 Notes.Rmd
milestone rep published

---
title: "Notes"
author: "Yigit Ozan Berk"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Course Title

Data Science Specialization SwiftKey Capstone

Course Instructor(s)

Jeff Leek
Roger Peng
Brian Caffo
You are a data scientist now

The goal of this data science specialization has been to give you the basic skills involved with being a data scientist. The goal of this capstone is to mimic the experience of being a data scientist. As a practicing data scientist it is entirely common to get a messy data set, a vague question, and very little instruction on exactly how to analyze the data. Our goal is to give you that same experience but with added support in the form of forums, discussion with instructors, feedback from SwiftKey and Coursera engineers, and a structured problem to solve. We hope that you will take advantage of the opportunity this project affords for you to demonstrate your skills and creativity.

Course Tasks

This course will be separated into 8 different tasks that cover the range of activities encountered by a practicing data scientist. They mirror many of the skills you have developed in the data science specialization. The tasks are:

Understanding the problem
Data acquisition and cleaning
Exploratory analysis
Statistical modeling
Predictive modeling
Creative exploration
Creating a data product
Creating a short slide deck pitching your product
You will hear about each of these tasks over the course of the capstone.

Assessements and Grading

To successfully complete the capstone project, you must receive a passing grade on all of the following assignments:

Quiz 1: Getting Started
Milestone Report: exploratory analysis of the data set + evaluation of at least three classmate submissions
Quiz 2: Natural Language Processing I
Quiz 3: Natural Language Processing II
Final Project: your data product and a presentation describing your final data product + evaluation of at least three classmate submissions
The quizzes will be standard multiple choice quizzes. The other components are graded by peer evaluation.

Your final grade will be calculated as follows

Quiz 1 - 5%
Milestone Report - 20%
Quiz 2 - 10%
Quiz 3 -10%
Final Project -55%
Course dataset

This is the training data to get you started that will be the basis for most of the capstone. You must download the data from the link below and not from external websites to start.

https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip
Later in the course you may use external data sets to augment your model as you see fit.

## Regular exp notes

whitespace word boundaries

sets of literals

beginnig and end of a line

alternatives (word1 or word2)

- Start of a line
"^i think"

- end of the line
"morning$"

- sets of characters
"[Bb][Uu][Ss][Hh]"
all case alternatives for the word in e.g. BUSH, BuSH, Bush, bush, bUSH

- all letters from a to z (uppercase or lower case)or numbers from 0 to 9, in the beginning of a line
"^[0-9][a-zA-Z]"

- when you use ^ at the beginning of a character class, it is a metacharacter indicating logical NOT operator.
"[^?.]$"  means = any lines NOT ending with a question mark or a period.

- "any" metacharacter
"9.11" means= 9 followed by any character followed by 11

- "or" metacharacter
"flood|fire|coldfire" will match flood or fire or coldfire

- you can combine
"^[Gg]ood|[Bb]ad" will match Good/good in the beginning of a line or Bad/bad anywhere

"^([Gg]ood|[Bb]ad)" will match Good/good or Bad/bad in the beginning of a line

- "optional" metacharacter
"[Gg]eorge( [Ww]\.)? [Bb]ush" will match where whitespace + [Ww] + . is optional

- backslash escapes metacharacter so . means any character, \. means period

- repetitions
"*" means repeat any number of times, including none
"+" means at least one

"(.*)" will match any character of any number

"[0-9]+ (.*)[0-9]+" will match, at least one number, followed by whitespace and any number of characters and at least one number again.

- you can quantify the number of repetitions (min and max) you want with the curly brackets{}

"[Bb]ush( +[^ ]+ +){1,5} debate" will match at least one space, followed by something that's not a space, followed by at least one space, and we want to see that between one and five times = we want to see something like space word space.

{m,n} means at least m but not more than n matches
{m} means exactly m matches
{m,} means at least m matches

- in most implementations of regular expressions, the parentheses not only limit the scope of alternatives divided by a "|", but also can be used to "remember" text matched by the subexpression enclosed
we refer to the matched text with \1, \2, etc..

" +([a-zA-Z]+) +\1 +" will match a space, followed by (at least one characters) followed by at least one space, followed by exactly the same match as before.
ex: "............. so so ......."

- * is greedy. it takes the longest possible string. you can use the ? to turn the greediness off.

packages to use

quanteda
readtext
spacyr

```{r}
install.packages("quanteda")
install.packages("readtext")
install.packages("spacyr")
install.packages("wordcloud")
```

# Loading in files using the readtext package
```{r}
require(readtext)

# Twitter json
dat_json <- readtext("~/Dropbox/QUANTESS/social media/zombies/tweets.json")
corp_twitter <- corpus(dat_json)
summary(corp_twitter, 5)
# generic json - needs a textfield specifier
dat_sotu <- readtext("~/Dropbox/QUANTESS/Manuscripts/collocations/Corpora/sotu/sotu.json",
                  textfield = "text")
summary(corpus(dat_sotu), 5)
# text file
dat_txtone <- readtext("~/Dropbox/QUANTESS/corpora/project_gutenberg/pg2701.txt", cache = FALSE)
summary(corpus(dat_txtone), 5)
# multiple text files
dat_txtmultiple1 <- readtext("~/Dropbox/QUANTESS/corpora/inaugural/*.txt", cache = FALSE)
summary(corpus(dat_txtmultiple1), 5)
# multiple text files with docvars from filenames
dat_txtmultiple2 <- readtext("~/Dropbox/QUANTESS/corpora/inaugural/*.txt",
                             docvarsfrom = "filenames", sep = "-",
                             docvarnames = c("Year", "President"))
summary(corpus(dat_txtmultiple2), 5)
# XML data
dat_xml <- readtext("~/Dropbox/QUANTESS/quanteda_working_files/xmlData/plant_catalog.xml",
                  textfield = "COMMON")
summary(corpus(dat_xml), 5)
# csv file
write.csv(data.frame(inaug_speech = texts(data_corpus_inaugural),
                     docvars(data_corpus_inaugural)),
          file = "/tmp/inaug_texts.csv", row.names = FALSE)
dat_csv <- readtext("/tmp/inaug_texts.csv", textfield = "inaug_speech")
summary(corpus(dat_csv), 5)
```

example

```{r}
ltcorpus = NULL
ltcorpus[1] = "SOS buy the book EOS"   
ltcorpus[2] = "SOS buy the book EOS"   
ltcorpus[3] = "SOS buy the book EOS"   
ltcorpus[4] = "SOS buy the book EOS"   
ltcorpus[5] = "SOS sell the book EOS"
ltcorpus[6] = "SOS buy the house EOS"
ltcorpus[7] = "SOS buy the house EOS"
ltcorpus[8] = "SOS paint the house EOS"
```

Step 1. i. Unigram, Bigram and Trigram counts
```{r}
## Returns a named vector of n-grams and their associated frequencies
## extracted from the character vector dat.
##
## ng - Defines the type of n-gram to be extracted: unigram if ng=1,
##      bigram if ng=2, trigram if n=3, etc.
## dat - Character vector from which we want to get n-gram counts.
## ignores - Character vector of words (features) to ignore from frequency table
## sort.by.ngram - sorts the return vector by the names
## sort.by.freq - sorts the return vector by frequency/count
getFreqs = function(dat, ng) {
        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
                      remove = stopwords("english"))
        dat.freq = docfreq(dat.dfm)
        return(dat.freq)
}

getNgramFreqs <- function(ng, dat, ignores=NULL,
                          sort.by.ngram=TRUE, sort.by.freq=FALSE) {
    # http://stackoverflow.com/questions/36629329/
    # how-do-i-keep-intra-word-periods-in-unigrams-r-quanteda
    if(is.null(ignores)) {
        dat.dfm <- dfm(dat, ngrams=ng, remove_punct = T, remove_numbers = T, 
                       remove = stopwords("english"), 
                       what = "fastestword")
    } else {
            dat.dfm <- dfm(dat, ngrams=ng, remove_punct = T, remove_numbers = T,
                           remove = stopwords("english"), 
                           ignoredFeatures = ignores, what = "fastestword")
    }
    rm(dat)
    # quanteda docfreq will get the document frequency of terms in the dfm
    ngram.freq <- docfreq(dat.dfm)
    if(sort.by.freq) { ngram.freq <- sort(ngram.freq, decreasing=TRUE) }
    if(sort.by.ngram) { ngram.freq <- ngram.freq[sort(names(ngram.freq))] }
    rm(dat.dfm)

    return(ngram.freq)
}

## Returns a 2 column data.table. The first column: ngram, contains all the
## unigrams, bigrams, or trigrams in the corpus depending on whether
## ng = 1, 2, or 3 respectively. The second column: freq, contains the
## frequency or count of the ngram found in linesCorpus.
##
## ng - Defines the type of n-gram to be extracted: unigram if ng=1,
##      bigram if ng=2, trigram if n=3, etc.
## linesCorpus - character vector: each element is a line from a corpus file
## prefixFilter - character vector: If not NULL, tells the function to return
##                only rows where the ngram column starts with prefixFilter.
##                If NULL, returns all the ngram and count rows.

getTables = function(dat, ng) {
        ngrams = getFreqs(dat = dat, ng = ng)
        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)
        return(ngrams_dt)
}

getNgramTables <- function(ng, linesCorpus, prefixFilter=NULL) {
    ngrams <- getNgramFreqs(ng, linesCorpus)
    ngrams_dt <- data.table(ngram=names(ngrams), freq=ngrams)
#    if(length(grep('^SOS', ngrams_dt$ngram)) > 0) {
 #       ngrams_dt <- ngrams_dt[-grep('^SOS', ngrams_dt$ngram),]
  #  }
   # if(!is.null(prefixFilter)) {
    #    regex <- sprintf('%s%s', '^', prefixFilter)
    #    ngrams_dt <- ngrams_dt[grep(regex, ngrams_dt$ngram),]
   # }
   # 
    return(ngrams_dt)
}

unigs <- getNgramTables(1, ltcorpus)
bigrs <- getNgramTables(2, ltcorpus)
trigs <- getNgramTables(3, ltcorpus)
unigs; bigrs; trigs
```

Step 1. ii. Selecting bigram and trigram discounts

For this example, we’ll use γ2=γ3=0.5
 for the purpose of illustration. In practice, these values would be obtained by cross-validation. A great treatment of cross-validation can be found in Chapter 5 of this (free) book [10] which are discussed by the authors in these three videos: [11], [12], and [13].

Step 2. Select Bigram Prefix of Word to be Predicted

For this example, we’ll select the bigram: sell the

Step 3. Calculate Probabilities of Words Completing Observed Trigrams

```{r}
gamma2 <- 0.5  # bigram discount
gamma3 <- 0.5  # trigram discount
bigPre <- 'sell_the'

## Returns a two column data.frame of observed trigrams that start with the
## bigram prefix (bigPre) in the first column named ngram and
## frequencies/counts in the second column named freq. If no observed trigrams
## that start with bigPre exist, an empty data.frame is returned.
##
## bigPre -  single-element char array of the form w2_w1 which are the first 
##           two words of the trigram we are predicting the tail word of
## trigrams - 2 column data.frame or data.table. The first column: ngram,
##            contains all the trigrams in the corpus. The second column:
##            freq, contains the frequency/count of each trigram.
getObsTrigs <- function(bigPre, trigrams) {
    trigs.winA <- data.frame(ngrams=vector(mode = 'character', length = 0),
                             freq=vector(mode = 'integer', length = 0))
    regex <- sprintf("%s%s%s", "^", bigPre, "_")
    trigram_indices <- grep(regex, trigrams$ngram)
    if(length(trigram_indices) > 0) {
        trigs.winA <- trigrams[trigram_indices, ]
    }

    return(trigs.winA)
}

## Returns a two column data.frame of observed trigrams that start with bigram
## prefix bigPre in the first column named ngram and the probabilities
## q_bo(w_i | w_i-2, w_i-1) in the second column named prob calculated from
## eqn 12. If no observed trigrams starting with bigPre exist, NULL is returned.
##
## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
##            contains all the observed trigrams that start with the bigram
##            prefix bigPre which we are attempting to the predict the next
##            word of in a give phrase. The second column: freq, contains the
##            frequency/count of each trigram.
## bigrs - 2 column data.frame or data.table. The first column: ngram,
##         contains all the bigrams in the corpus. The second column:
##         freq, contains the frequency/count of each bigram.
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## triDisc - amount to discount observed trigrams
getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
    if(nrow(obsTrigs) < 1) return(NULL)
    obsCount <- filter(bigrs, ngram==bigPre)$freq[1]
    obsTrigProbs <- mutate(obsTrigs, freq=((freq - triDisc) / obsCount))
    colnames(obsTrigProbs) <- c("ngram", "prob")

    return(obsTrigProbs)
}

obs_trigs <- getObsTrigs(bigPre, trigs)  # get trigrams and counts
# convert counts to probabilities
qbo_obs_trigrams <- getObsTriProbs(obs_trigs, bigrs, bigPre, gamma3)
qbo_obs_trigrams
```

Step 4. Calculate Probabilities of Words Completing Unobserved Trigrams

This is the the most complex step as it involves backing off to the bigram level. Here is a breakdown of the sub-steps for these calculations:

Find all the words that complete unobserved trigrams. These are the words in the set w∈(wi−2,wi−1)
 described earlier.
Calculate α(wi−1)
 from equation 14.
Calculate qBO
 for each bigram in the denominator of equation 17. using equation 10. if the bigram is observed or equation 16. if it is unobserved.
Calculate α(wi−2,wi−1)
 from equation 15.
Calculate qBO(wi|wi−2,wi−1)
 for each wi
 from equation 17.
Step 4. i. Find unobserved trigram tail words:

```{r}
## Returns a character vector which are the tail words of unobserved trigrams
## that start with the first two words of obsTrigs (aka the bigram prefix).
## These are the words w in the set B(w_i-2, w_i-1) as defined in the section
## describing the details of equation 17.
##
## obsTrigs - character vector of observed trigrams delimited by _ of the form:
##            w3_w2_w1 where w3_w2 is the bigram prefix
## unigs - 2 column data.frame of all the unigrams in the corpus:
##         ngram = unigram
##         freq = frequency/count of each unigram
getUnobsTrigTails <- function(obsTrigs, unigs) {
    obs_trig_tails <- str_split_fixed(obsTrigs, "_", 3)[, 3]
    unobs_trig_tails <- unigs[!(unigs$ngram %in% obs_trig_tails), ]$ngram
    return(unobs_trig_tails)
}

unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
unobs_trig_tails
```

Step 4. ii. Calculate discounted probability mass at the bigram level α(wi−1)
:

The code below implements equation 14. to calculate α(wi−1)

```{r}
## Returns the total probability mass discounted from all observed bigrams
## calculated from equation 14.  This is the amount of probability mass which
## is redistributed to UNOBSERVED bigrams. If no bigrams starting with
## unigram$ngram[1] exist, 0 is returned.
##
## unigram - single row, 2 column frequency table. The first column: ngram,
##           contains the w_i-1 unigram (2nd word of the bigram prefix). The
##           second column: freq, contains the frequency/count of this unigram.
## bigrams - 2 column data.frame or data.table. The first column: ngram,
##           contains all the bigrams in the corpus. The second column:
##           freq, contains the frequency or count of each bigram.
## bigDisc - amount to discount observed bigrams
getAlphaBigram <- function(unigram, bigrams, bigDisc=0.5) {
    # get all bigrams that start with unigram
    regex <- sprintf("%s%s%s", "^", unigram$ngram[1], "_")
    bigsThatStartWithUnig <- bigrams[grep(regex, bigrams$ngram),]
    if(nrow(bigsThatStartWithUnig) < 1) return(0)
    alphaBi <- 1 - (sum(bigsThatStartWithUnig$freq - bigDisc) / unigram$freq)

    return(alphaBi)
}

unig <- str_split(bigPre, "_")[[1]][2]
unig <- unigs[unigs$ngram == unig,]
alpha_big <- getAlphaBigram(unig, bigrs, gamma2)
alpha_big
```

Step 4. iii. Calculate backed off probabilities qBO
 for bigrams

The code below implements equation 10. to calculate qBO(wi|wi−1)
 for observed bigrams and equation 16. for unobserved bigrams:

```{r}
## Returns a character vector of backed off bigrams of the form w2_w1. These 
## are all the (w_i-1, w) bigrams where w_i-1 is the tail word of the bigram
## prefix bigPre and w are the tail words of unobserved bigrams that start with
## w_i-1.
##
## bigPre - single-element char array of the form w2_w1 which are first two
##          words of the trigram we are predicting the tail word of
## unobsTrigTails - character vector that are tail words of unobserved trigrams
getBoBigrams <- function(bigPre, unobsTrigTails) {
    w_i_minus1 <- str_split(bigPre, "_")[[1]][2]
    boBigrams <- paste(w_i_minus1, unobsTrigTails, sep = "_")
    return(boBigrams)
}

## Returns a two column data.frame of backed-off bigrams in the first column
## named ngram and their frequency/counts in the second column named freq.
## 
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## unobsTrigTails - character vector that are tail words of unobserved trigrams
## bigrs - 2 column data.frame or data.table. The first column: ngram,
##         contains all the bigrams in the corpus. The second column:
##         freq, contains the frequency/count of each bigram.
getObsBoBigrams <- function(bigPre, unobsTrigTails, bigrs) {
    boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
    obs_bo_bigrams <- bigrs[bigrs$ngram %in% boBigrams, ]
    return(obs_bo_bigrams)
}

## Returns a character vector of backed-off bigrams which are unobserved.
##
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## unobsTrigTails - character vector that are tail words of unobserved trigrams
## obsBoBigram - data.frame which contains the observed bigrams in a column
##               named ngram
getUnobsBoBigrams <- function(bigPre, unobsTrigTails, obsBoBigram) {
    boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
    unobs_bigs <- boBigrams[!(boBigrams %in% obsBoBigram$ngram)]
    return(unobs_bigs)
}

## Returns a dataframe of 2 columns: ngram and probs.  Values in the ngram
## column are bigrams of the form: w2_w1 which are observed as the last
## two words in unobserved trigrams.  The values in the prob column are
## q_bo(w1 | w2) calculated from from equation 10.
##
## obsBoBigrams - a dataframe with 2 columns: ngram and freq. The ngram column
##                contains bigrams of the form w1_w2 which are observed bigrams
##                that are the last 2 words of unobserved trigrams (i.e. "backed
##                off" bigrams). The freq column contains integers that are
##                the counts of these observed bigrams in the corpus.
## unigs - 2 column data.frame of all the unigrams in the corpus:
##         ngram = unigram
##         freq = frequency/count of each unigram
## bigDisc - amount to discount observed bigrams
getObsBigProbs <- function(obsBoBigrams, unigs, bigDisc=0.5) {
    first_words <- str_split_fixed(obsBoBigrams$ngram, "_", 2)[, 1]
    first_word_freqs <- unigs[unigs$ngram %in% first_words, ]
    obsBigProbs <- (obsBoBigrams$freq - bigDisc) / first_word_freqs$freq
    obsBigProbs <- data.frame(ngram=obsBoBigrams$ngram, prob=obsBigProbs)

    return(obsBigProbs)
}

## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
## column are unobserved bigrams of the form: w2_w1.  The values in the prob
## column are the backed off probability estimates q_bo(w1 | w2) calculated
## from from equation 16.
##
## unobsBoBigrams - character vector of unobserved backed off bigrams
## unigs - 2 column data.frame of all the unigrams in the corpus:
##         ngram = unigram
##         freq = frequency/count of each unigram
## alphaBig - total discounted probability mass at the bigram level
getQboUnobsBigrams <- function(unobsBoBigrams, unigs, alphaBig) {
    # get the unobserved bigram tails
    qboUnobsBigs <- str_split_fixed(unobsBoBigrams, "_", 2)[, 2]
    w_in_Aw_iminus1 <- unigs[!(unigs$ngram %in% qboUnobsBigs), ]
    # convert to data.frame with counts
    qboUnobsBigs <- unigs[unigs$ngram %in% qboUnobsBigs, ]
    denom <- sum(qboUnobsBigs$freq)
    # converts counts to probabilities
    qboUnobsBigs <- data.frame(ngram=unobsBoBigrams,
                               prob=(alphaBig * qboUnobsBigs$freq / denom))

    return(qboUnobsBigs)
}

bo_bigrams <- getBoBigrams(bigPre, unobs_trig_tails)  # get backed off bigrams
# separate bigrams which use eqn 10 and those that use 16
obs_bo_bigrams <- getObsBoBigrams(bigPre, unobs_trig_tails, bigrs)
unobs_bo_bigrams <- getUnobsBoBigrams(bigPre, unobs_trig_tails, obs_bo_bigrams)
# unobs_bo_bigrams = c("the_buy", "the_EOS", "the_paint", "the_sell", "the_the")
# calc obs'd bigram prob's from eqn 10
qbo_obs_bigrams <- getObsBigProbs(obs_bo_bigrams, unigs, gamma2) #ngram     probs
# calc alpha_big & unobs'd bigram prob's from eqn 16             #the_house 0.3125
unig <- str_split(bigPre, "_")[[1]][2]
unig <- unigs[unigs$ngram == unig,]
# distrib discounted bigram prob mass to unobs bigrams in prop to unigram ML
qbo_unobs_bigrams <- getQboUnobsBigrams(unobs_bo_bigrams, unigs, alpha_big)
qbo_bigrams <- rbind(qbo_obs_bigrams, qbo_unobs_bigrams)
qbo_bigrams
```

Checking the Bigram Calculations

Before doing the final calculations for the unobserved trigrams, let’s do a simple check on our calculations at the bigram level. In the previous table, all the bigrams except the_house are unobserved which means that if we sum all the unobserved bigram probabilities, we should get the total bigram discount which is α(wi−1)
. As we see below, this looks like it checks out.

```{r}
unobs <- qbo_bigrams[-1,]
sum(unobs$prob)
```

Step 4. iv. Calculate discounted probability mass at the trigram level α(wi−2,wi−1)

The getAlphaTrigram function shown below implements equation 15. to compute the trigram discount. Here we use it to compute the trigram discount for qBO(house|sell,the)
:

```{r}
## Returns the total probability mass discounted from all observed trigrams.
## calculated from equation 14. This is the amount of probability mass which is
## redistributed to UNOBSERVED trigrams. If no trigrams starting with
## bigram$ngram[1] exist, 1 is returned.
##
## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
##            contains all the observed trigrams that start with the bigram
##            prefix we are attempting to the predict the next word of. The 
##            second column: freq, contains the frequency/count of each trigram.
## bigram - single row frequency table where the first col: ngram, is the bigram
##          which are the first two words of unobserved trigrams we want to
##          estimate probabilities of (same as bigPre in other functions listed
##          prior) delimited with '_'. The second column: freq, is the
##          frequency/count of the bigram listed in the ngram column.
## triDisc - amount to discount observed trigrams
getAlphaTrigram <- function(obsTrigs, bigram, triDisc=0.5) {
    if(nrow(obsTrigs) < 1) return(1)
    alphaTri <- 1 - sum((obsTrigs$freq - triDisc) / bigram$freq[1])

    return(alphaTri)
}

bigram <- bigrs[bigrs$ngram %in% bigPre, ]
alpha_trig <- getAlphaTrigram(obs_trigs, bigram, gamma3)
alpha_trig
```

Step 4. v. Calculate unobserved trigram probabilities qBO(wi|wi−2,wi−1)
:

```{r}
## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
## column are unobserved trigrams of the form: w3_w2_w1.  The values in the prob
## column are q_bo(w1 | w3, w2) calculated from equation 17.
##
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## qboObsBigrams - 2 column data.frame with the following columns -
##                 ngram: observed bigrams of the form w2_w1
##                 probs: the probability estimate for observed bigrams:
##                        qbo(w1 | w2) calc'd from equation 10.
## qboUnobsBigrams - 2 column data.frame with the following columns -
##                   ngram: unobserved bigrams of the form w2_w1
##                   probs: the probability estimate for unobserved bigrams
##                          qbo(w1 | w2) calc'd from equation 16.
## alphaTrig - total discounted probability mass at the trigram level
getUnobsTriProbs <- function(bigPre, qboObsBigrams,
                             qboUnobsBigrams, alphaTrig) {
    qboBigrams <- rbind(qboObsBigrams, qboUnobsBigrams)
    qboBigrams <- qboBigrams[order(-qboBigrams$prob), ]
    sumQboBigs <- sum(qboBigrams$prob)
    first_bigPre_word <- str_split(bigPre, "_")[[1]][1]
    unobsTrigNgrams <- paste(first_bigPre_word, qboBigrams$ngram, sep="_")
    unobsTrigProbs <- alphaTrig * qboBigrams$prob / sumQboBigs
    unobsTrigDf <- data.frame(ngram=unobsTrigNgrams, prob=unobsTrigProbs)

    return(unobsTrigDf)
}

qbo_unobs_trigrams <- getUnobsTriProbs(bigPre, qbo_obs_bigrams,
                                       qbo_unobs_bigrams, alpha_trig)
qbo_unobs_trigrams
```

Step 5. Select wi
 with the highest qBO(wi|wi−2,wi−1)

We’ve done all the calculations required to make our prediction. These are summarized in the table below:

```{r}
getPredictionMsg <- function(qbo_trigs) {
    # pull off tail word of highest prob trigram
    prediction <- str_split(qbo_trigs$ngram[1], "_")[[1]][3]
    result <- sprintf("%s%s%s%.4f", "highest prob prediction is >>> ", prediction,
                      " <<< which has probability = ", qbo_trigs$prob[1])
    return(result)
}

qbo_trigrams <- rbind(qbo_obs_trigrams, qbo_unobs_trigrams)
qbo_trigrams <- qbo_trigrams[order(-qbo_trigrams$prob), ]  # sort by desc prob
out_msg <- getPredictionMsg(qbo_trigrams)
out_msg
```