### 2019-11-09 draft1.Rmd

---
title: "SwiftKey Capstone Proj. Milestone Report"
author: "Yigit Ozan Berk"
date: "11/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initiation
```{r}
library(quanteda)
library(readtext)
library(spacyr)
library(ggplot2)
library(wordcloud)
library(data.table)
```

# Downloading Data

The dataset can be downloaded from a link given in the course website. [https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip]

The unzipped file contains a directory called final, then a subdirectory called en_US, which contains the texts.

There are 3 text files.
* en_US.blogs.txt - text from blog posts
* en_US.news.txt - text from news articles
* en_US.twitter.txt - tweets on Twitter

# Introduction

The goal here to display the initial data exploration, and show I am on the track to extablishing my algorithm.

```{r}
numwords <- system("wc -w *.txt", intern=TRUE)  # intern=TRUE to return output  
numlines <- system("wc -l *.txt", intern=TRUE)
numbytes <- system("wc -c *.txt", intern=TRUE)

# number of words for each dataset
blog.numwords <- as.numeric(gsub('[^0-9]', '', numwords[1]))
news.numwords <- as.numeric(gsub('[^0-9]', '', numwords[2]))
twit.numwords <- as.numeric(gsub('[^0-9]', '', numwords[3]))

# number of lines for each dataset
blog.numlines <- as.numeric(gsub('[^0-9]', '', numlines[1]))
news.numlines <- as.numeric(gsub('[^0-9]', '', numlines[2]))
twit.numlines <- as.numeric(gsub('[^0-9]', '', numlines[3]))

# number of bytes for each dataset
blog.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[1]))
news.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[2]))
twit.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[3]))

words = rbind(blog.numwords, news.numwords, twit.numwords)
lines = rbind(blog.numlines, news.numlines, twit.numlines)
bytes = rbind(blog.numbytes, news.numbytes, twit.numbytes)

data.frame(words = words, lines= lines, Mb = bytes/1000000, 
           row.names = c("blog", "news", "twit"))
```

For memory storage limitations, only %15 random entries from each dataset is used for calculations.

Twitter text
```{r}
con <- file("./en_US.twitter.txt", "r") 
twit = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(1)
x = sample(2360148, 350000, replace = F)

train = twit[x]

TW <- corpus(train)
rm(con, train, twit, x)
summary(TW, 5)
```

Blog text
```{r}
con <- file("./en_US.blogs.txt", "r") 
blog = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(12345)
x = sample(899288, 135000, replace = F)

train = blog[x]

BG <- corpus(train)
rm(con, train, blog, x)
summary(BG, 5)
```

News text

```{r}
# 1010242
con <- file("./en_US.news.txt", "r") 
news = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(10000)
x = sample(1010242, 150000, replace = F)

train = news[x]

NS <- corpus(train)
rm(con, train, news, x)
summary(NS, 5)
```

# Visualize Word Frequency

```{r}
BG.uni = getTables(dat = BG, ng = 1)
set.seed(100)
textplot_wordcloud(BG.uni, min_count = 3000, random_order = FALSE,
                   rotation = .25,
                   color = RColorBrewer::brewer.pal(8, "Dark2"))
```

1. create frequency tables and n-grams
```{r}
getFreqs = function(dat, ng) {
        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
                      remove = stopwords("english"))
        dat.freq = docfreq(dat.dfm)
        dat.freq = dat.freq[sort(names(dat.freq))] 
        return(dat.freq)
}

getTables = function(dat, ng) {
        ngrams = getFreqs(dat = dat, ng = ng)
        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)
        return(ngrams_dt)
}
```

```{r}
gamma2 <- 0.5  # bigram discount
gamma3 <- 0.5  # trigram discount

bigPre <- 'sell_the'

getObsTrigs <- function(bigPre, trigrams) {
    trigs.winA <- data.frame(ngrams=vector(mode = 'character', length = 0),
                             freq=vector(mode = 'integer', length = 0))
    regex <- sprintf("%s%s%s", "^", bigPre, "_")
    trigram_indices <- grep(regex, trigrams$ngram)
    if(length(trigram_indices) > 0) {
        trigs.winA <- trigrams[trigram_indices, ]
    }

    return(trigs.winA)
}

getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
    if(nrow(obsTrigs) < 1) return(NULL)
    obsCount <- filter(bigrs, ngram==bigPre)$freq[1]
    obsTrigProbs <- mutate(obsTrigs, freq=((freq - triDisc) / obsCount))
    colnames(obsTrigProbs) <- c("ngram", "prob")

    return(obsTrigProbs)
}

obs_trigs <- getObsTrigs(bigPre, trigs)  # get trigrams and counts
# convert counts to probabilities
qbo_obs_trigrams <- getObsTriProbs(obs_trigs, bigrs, bigPre, gamma3)
qbo_obs_trigrams
```

```{r}
getUnobsTrigTails <- function(obsTrigs, unigs) {
    obs_trig_tails <- str_split_fixed(obsTrigs, "_", 3)[, 3]
    unobs_trig_tails <- unigs[!(unigs$ngram %in% obs_trig_tails), ]$ngram
    return(unobs_trig_tails)
}

unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
unobs_trig_tails
```