### 2019-11-16 al_stup.Rmd
stupid backoff

---
title: "stupid backoff"
author: "Yigit Ozan Berk"
date: "11/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(quanteda)
library(dplyr)
library(stringr)
library(readtext)
library(spacyr)
library(ggplot2)
library(wordcloud)
library(data.table)
```

# Downloading Data

The dataset can be downloaded from a link given in the course website. [https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip]

The unzipped file contains a directory called final, then a subdirectory called en_US, which contains the texts.

There are 3 text files.
* en_US.blogs.txt - text from blog posts
* en_US.news.txt - text from news articles
* en_US.twitter.txt - tweets on Twitter

# Introduction

The goal here to display the initial data exploration, and show I am on the track to extablishing my algorithm.

```{r}
numwords <- system("wc -w *.txt", intern=TRUE)  # intern=TRUE to return output  
numlines <- system("wc -l *.txt", intern=TRUE)
numbytes <- system("wc -c *.txt", intern=TRUE)

# number of words for each dataset
blog.numwords <- as.numeric(gsub('[^0-9]', '', numwords[1]))
news.numwords <- as.numeric(gsub('[^0-9]', '', numwords[2]))
twit.numwords <- as.numeric(gsub('[^0-9]', '', numwords[3]))

# number of lines for each dataset
blog.numlines <- as.numeric(gsub('[^0-9]', '', numlines[1]))
news.numlines <- as.numeric(gsub('[^0-9]', '', numlines[2]))
twit.numlines <- as.numeric(gsub('[^0-9]', '', numlines[3]))

# number of bytes for each dataset
blog.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[1]))
news.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[2]))
twit.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[3]))

words = rbind(blog.numwords, news.numwords, twit.numwords)
lines = rbind(blog.numlines, news.numlines, twit.numlines)
bytes = rbind(blog.numbytes, news.numbytes, twit.numbytes)

data.frame(words = words, lines= lines, Mb = bytes/1000000, 
           row.names = c("blog", "news", "twit"))
```

For memory storage limitations, only %15 random entries from each dataset is used for calculations.

Twitter text
```{r}
con <- file("./en_US.twitter.txt", "r") 
twit = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(1)
x = sample(2360148, 350000, replace = F)

train = twit[x]

TW <- corpus(train)
rm(con, train, twit, x)
summary(TW, 5)
```

Blog text
```{r}
con <- file("./en_US.blogs.txt", "r") 
blog = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(12345)
x = sample(899288, 135000, replace = F)

train = blog[x]

BG <- corpus(train)
rm(con, train, blog, x)
summary(BG, 5)
```

News text

```{r}
# 1010242
con <- file("./en_US.news.txt", "r") 
news = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(10000)
x = sample(1010242, 150000, replace = F)

train = news[x]

NS <- corpus(train)
rm(con, train, news, x)
summary(NS, 5)
```

# Visualize Word Frequency

```{r}
BG.uni = getTables(dat = BG, ng = 1)
set.seed(100)
textplot_wordcloud(BG.uni, min_count = 3000, random_order = FALSE,
                   rotation = .25,
                   color = RColorBrewer::brewer.pal(8, "Dark2"))
```

1. create frequency tables and n-grams
```{r}
getFreqs = function(dat, ng) {
        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
                      remove = stopwords("english"))
        dat.freq = docfreq(dat.dfm)
        dat.freq = dat.freq[sort(names(dat.freq))] 
        return(dat.freq)
}

getTables = function(dat, ng) {
        ngrams = getFreqs(dat = dat, ng = ng)
        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)
        return(ngrams_dt)
}

```

unigram, bigram, and trigrams

```{r}
unigs = getTables(BG, 1)
bigrs = getTables(BG, 2)
trigs = getTables(BG, 3)
```

# Algorithm

Probs of words completing OBSERVED trigrams
```{r}
bigPre <- 'is_book'

## Returns a two column data.frame of observed trigrams that start with the
## bigram prefix (bigPre) in the first column named ngram and
## frequencies/counts in the second column named freq. If no observed trigrams
## that start with bigPre exist, an empty data.frame is returned.
##
## bigPre -  single-element char array of the form w2_w1 which are the first 
##           two words of the trigram we are predicting the tail word of
## trigrams - 2 column data.table. The first column: ngram,
##            contains all the trigrams in the corpus. The second column:
##            freq, contains the frequency/count of each trigram.
getODTrigs = function(bigPre, trigrams) {
    tri.obs = data.frame(ngrams = vector(mode = 'character', length = 0),
                          freq = vector(mode = 'integer', length = 0))
    regex = sprintf("%s%s%s", "^", bigPre, "_")
    indices = grep(regex, trigrams$ngram)
    if(length(indices > 0)) {
        tri.obs = trigrams[indices, ]
    }
    return(tri.obs)
}

getODBigs = function(bigPre, bigrams) {
       big.obs = data.frame(ngrams = vector(mode = 'character', length = 0),
                            freq = vector(mode = 'integer', length = 0))
       w_i_1 <- str_split(bigPre, "_")[[1]][2]
       regex = sprintf("%s%s%s", "^", w_i_1, "_")
       indices = grep(regex, bigrams$ngram)
       if(length(indices > 0)) {
               big.obs = bigrams[indices, ]
       }
       return(big.obs)
}

getODUnigs = function(bigPre, unigrams) {

}

getTrigProbs = function(bigPre, trigrams, bigrams) {

}

S3 = getODTrigs(bigPre, trigs) # get observed trigrams and counts
S2 = getODBigs(bigPre, bigrs) # get observed bigrams and counts

```